Categories of Optics

### 2018-06-26 paper.tex
Explain Folds and fix some typos

\newcommand{\Act}{\mathbf{Act}}
\newcommand{\Optic}{\mathbf{Optic}}
\newcommand{\Lawful}{\mathbf{Lawful}}
%\newcommand{\SemiOptic}{\mathbf{SemiOptic}}
\newcommand{\Lens}{\mathbf{Lens}}
\newcommand{\Prism}{\mathbf{Prism}}
\newcommand{\Setter}{\mathbf{Setter}}
\newcommand{\Traversal}{\mathbf{Traversal}}
\newcommand{\Getter}{\mathbf{Getter}}
\newcommand{\Review}{\mathbf{Review}}

@@

\begin{lemma}[Ninja Yoneda Lemma/Yoneda Reduction]
For every functor $K : \C^\op \to \Set$ and $H : \C \to \Set$, we have the following natural isomorphisms:
\begin{align*}
K &\cong \int^C KC \times \C(-,C) &
K &\cong \int_C (KC)^{\C(C,-)} \\
H &\cong \int^C HC \times \C(C,-)  &
H &\cong \int_C (HC)^{\C(-,C)}
\end{align*}
where the components of the isomorphisms at $X$ are given by inclusion with the identity morphism $\C(X, X)$ on the left, and projection onto the identity morphism on the right.
\qed
\end{lemma}

@@

Because this coend takes place in $\Set$, we can use Lemma \ref{lemma:calculate-coend} to describe the set $\Optic_\otimes((S, S'), (A, A'))$ explicitly. It is the set of pairs $(l, r)$, where $l : S \to M \otimes A$ and $r : M \otimes A' \to S'$, quotiented by the relation
\begin{align*}
  ((f \otimes A) l, r) \sim (l, r (f \otimes A'))
\end{align*}
for any $l : S \to M \otimes A$, $r : N \otimes A' \to S'$ and $f : M \to N$.

@@

  To check extranaturality of $\varepsilon$, suppose we have a dualisable optic $\iota(f, g) : (S, S') \hto (T, T')$, so $f : S \to T$ and $g : T' \to S'$. Luckily here all the switching in the definitions scancels out! Extranaturality is witnessed by the equality of the string diagrams:
  \begin{center}
    \input{diagrams/counit-extranatural-left.tikz}
    \qquad \raisebox{1.5cm}{$=$} \qquad
    \input{diagrams/counit-extranatural-right.tikz}
  \end{center}
  For symmetry:
  \begin{center}
    \input{diagrams/counit-symmetry-left.tikz}
    \qquad \raisebox{1.5cm}{$=$} \qquad
    \input{diagrams/counit-symmetry-right.tikz}
  \end{center}
  And for monoidality there is essentially nothing to do in the graphical calculus. 
  \begin{center}
    \input{diagrams/counit-monoidal-left.tikz}
    \qquad \raisebox{2cm}{$=$} \qquad
    \input{diagrams/counit-monoidal-right.tikz}
  \end{center}

@@

  And it also preserves the counits:
  \begin{align*}
    F(\varepsilon_{(S, S')})
    &= F(c_{S \otimes S'}) \\
    &= F(\rep{\rho_S^{-1}}{\rho_{S'}}) \\
    &= \rep{\phi^{-1}_{S,I} (F \rho_S^{-1})}{(F \rho_{S'}) \phi_{S',I}} \\
    &= \rep{(FS \otimes \phi_I^{-1}) \phi^{-1}_{S,I} (F \rho_S^{-1})}{(F \rho_{S'}) \phi_{S',I} (FS \otimes \phi_I)} \\
    &= \rep{\rho_{FS}^{-1}}{\rho_{FS'} } \\
    &= \varepsilon_{(FS, FS')}
  \end{align*}

@@

\begin{definition}
Given two categories equipped with monoidal actions $(\M, \C)$ and $(\N, D)$, a \emph{morphism of actions} is a monoidal functor $F^\bullet : \M \to N$ and a functor $F : \C \to \D$ that commutes with the actions, in the sense that there exist natural isomorphisms
  \begin{align*}
  \phi &: F(I A) \to (F^\bullet I)(F_2A) \\
  \phi_{M,A} &: F(MA) \to (F^\bullet M)(F A) 
  \end{align*}
satisfying conditions analogous to those for a monoidal functor.
\end{definition}

@@

  Suppose we have two lawful optics $\rep{l}{r} : R \hto S$ and $\rep{l'}{r'} : S \hto A$, where $l : R \to MS$ and $l' : S \to NA$. We must show that $\rep{(Ml')l}{r (Mr') }$ is also a lawful optic. The first equation is straightforward:
  \begin{align*}
    \outside(\rep{l; (Ml')}{(Mr') ; r  })
    &= l ; Ml' ; Mr' ; r \\
    &= l ; M(l'r') ; r \\
    &= l ; M\id_{NA} ; r \\
    &= l ; r \\
    &= \id_S
  \end{align*}

  For the second, we must show
  \[ \repthree{ l;(Ml')}{(Mr'); r;l;(Ml')}{(Mr') ; r } = \repthree{l;(Ml')}{\id_{MNA}}{(Mr') ; r }. \]

@@

  So consider the generating relation $\repthree{ l;\phi_S}{c}{r} = \repthree{l}{\phi_S ; c}{r}$ in \[\int^{M_1, M_2 \in \M} \C(R, M_1 S) \times \C(M_1 S, M_2 S) \times \C(M_2 S, R).\]
  This implies that
  \begin{align*}
    \repthree{l;\phi_S;(Ml')}{(Mr'); c ;(Ml')}{(Mr') ; r}
    &= \repthree{l;(Ml');\phi_{NA}}{(Mr'); c ;(Ml')}{(Mr') ; r} && \text{(functoriality)} \\
    &= \repthree{l;(Ml')}{\phi_{NA};(Mr'); c ;(Ml')}{(Mr') ; r} && \text{(coend relation)} \\
    &= \repthree{l;(Ml')}{(Mr');\phi_{S}; c ;(Ml')}{(Mr') ; r} && \text{(functoriality)}
  \end{align*}
  And similarly for the other generating relation, $\repthree{l}{c;\phi_S}{r } = \repthree{l}{c}{\phi_S; r }$.

  Therefore, the chain of relations proving $\repthree{l}{r;l}{r } = \repthree{l}{\id_{MA}}{r }$ can be replicated to show
  \begin{align*}
    \repthree{l;(Ml')}{(Mr');r;l;(Ml')}{(Mr') ; r } &= \repthree{l;(Ml')}{(Mr');\id_{MA};(Ml')}{(Mr') ; r },
  \end{align*}
  which is equal to $\repthree{l;(Ml')}{M(r'l')}{(Mr') ; r }$.

  Similarly, a generating relation $\repthree{l';\psi_A}{c'}{r' } = \repthree{l'}{\psi_A ; c'}{r' }$ in \[\int^{N_1, N_2 \in \M} \C(S, N_1 A) \times \C(N_1 A, N_2 A) \times \C(M_2 A, S)\]
  implies that
  \begin{align*}
    \repthree{l;M(l';\psi_A)}{Mc'}{(Mr') ; r }
    &= \repthree{l;Ml';M\psi_A}{Mc'}{(Mr') ; r } \\
    &= \repthree{l;Ml'}{M\psi_A ; Mc'}{(Mr') ; r } \\
    &= \repthree{l;Ml'}{M(\psi_A;c')}{(Mr') ; r }
  \end{align*}
  So again we can replicate the chain of relations proving $\repthree{l'}{r';l'}{r' } = \repthree{l'}{\id_{MA}}{r' }$ to show that
  \[\repthree{l;(Ml')}{M(r'l')}{(Mr') ; r } = \repthree{l;(Ml')}{\id_{MNA}}{(Mr') ; r } \]

@@

\begin{proposition}
  Suppose $F : (\M, \C) \to (\N, \D)$ is a morphism of actions. Then $\Optic(F) : \Optic_\M \to \Optic_\N$ restricts to a functor $\Lawful_\M \to \Lawful_\N$.
\end{proposition}
\begin{proof}
  If $p = \rep{l}{r}$ is lawful, then verifying the first equation is easy:
  \begin{align*}
  \outside(\Optic(F)(\rep{l}{r})) 
  &= \outside(\rep{(Fl);\phi^{-1}_{M,A} }{\phi_{M,A'};(Fr) }) \\
  &= (Fl);\phi^{-1}_{M,A};\phi_{M,A'};(Fr)\\
  &= (Fl);(F_2r)\\
  &= \id_{FS}
  \end{align*}
  where $\phi_{M,A'} : (F^\bullet M)(FA) \to F(MA)$ is the structure map for $F$ commuting with the actions.

@@

\begin{proposition}
  Let $\rep{l}{r } : S \hto A$ be an optic. If $l;r = \id_S$ and $r;l = \phi A$ for some $\phi : M \to M$ in $\M$, then $\rep{l}{r }$ is lawful.
\end{proposition}
\begin{proof}
  The $\outside(\rep{l}{r }) = l;r = \id_S$ is exactly the first condition. And for the second, we verify:
  \begin{align*}
    \twice(\rep{l}{r })
    &= \repthree{l}{r;l}{r } \\
    &= \repthree{l}{\phi A}{r } && \text{($r;l = \phi A$)}\\
    &= \repthree{l (\phi A)}{\id_{MA}}{r } && \text{(coend relation)} \\
    &= \repthree{l;r;l}{\id_{MA}}{r } && \text{($r;l = \phi A$ again)}\\
    &= \repthree{l}{\id_{MA}}{r }  && \text{($l;r = \id_S$)}\\
    &= \once(\rep{l}{r })
  \end{align*}
\end{proof}

@@

Let $\inside : \Optic_\M(S, A) \to \int^{M \in \M} \C(M A, M A)$ be the map induced by $\inside(\rep{l}{r }) = \langle r ; l \rangle$. We might ask that instead of requiring $r;l = \phi A$ exactly, we have $\langle r ; l \rangle = \langle \phi A \rangle$ in $\int^{M \in \M} \C(M A, M A)$. In fact, this is equivalent:

\begin{proposition}\label{prop-onthenose}
  Suppose $p : S \hto A$ satisfies $\outside(p) = \id_S$ and $\inside(p) = \langle \phi A \rangle$. Then there exists a representative $\rep{l}{r }$ such that $l ; r = \psi A$ on the nose for some (possibly different) $\psi : M \to M$.
\end{proposition}
\begin{proof}
  The generating relation for $\int^{M \in \M} \C(M A, M A)$ is 
  \[ \langle f \phi_A \rangle = \langle \phi_A f \rangle \] 
  whenever $f : N A \to M A$ and $\phi : M \to N$. This relation $f \phi_A \rightsquigarrow \phi_A f$ is not likely to be symmetric or transitive in general.

@@

  We can now inductively apply the above argument to the shorter chain \[l'r' = u_1^2 \leftrightsquigarrow \dots \leftrightsquigarrow u_n^2 \leftrightsquigarrow (\psi A)^2 = \psi^2 A,\] obtained by squaring each morphism in the original chain, until we are left with a representative $\rep{l^*}{r^*}$ such that $r^*;l^* = \psi^N_A$, for some $N>0$. This pair $\rep{l^*}{r^* }$ is the required representative.
\end{proof}

@@

This last set is the set of concrete lenses $\conc_\times((S, S'), (A, A'))$. Explicitly this isomorphism states that, given an optic $\rep{l}{r } : (S, S') \hto (A, A')$, the corresponding element of $\conc_\times((S, S'), (A, A'))$ is the pair $(\fget, \fput)$, where $\fget = \pi_2 l$ and $\fput = r (\pi_1 l \times A)$. In the other direction, given $(\fget, \fput)$, the corresponding optic has representative $\rep{[\id_S, \fget]}{\fput}$.

@@

\begin{proposition}\label{prop-OpticImpliesLensLaws}
  A concrete lens given by $\fget$ and $\fput$ is lawful (in our sense) if and it obeys the three lens laws.
\end{proposition}
\begin{proof}
  We begin by giving
  \[ \int^{M_1, M_2 \in \C} \C(S, M_1 \times A) \times \C(M_1 \times A, M_2 \times A) \times \C(M_2 \times A, S)\]
  the same treatment as $\Optic_\times(S, A)$. By the same sorts of arguments as above \todo{is this worth writing out?}, this is isomorphic to
  \[\conctwice_\times(S, A) := \C(S, A) \times \C(S \times A, A) \times \C(S \times A \times A, S)\]
  with the isomorphism given by
  \begin{align*}
    \Phi(\repthree{l}{c}{r }) = (&\pi_2 l, \\
                                   &\pi_2 c (\pi_1 l \times A), \\
                                   &r(\pi_1 c (\pi_1 l \times A) \times A))
  \end{align*}
  Now suppose we are given a lens $(\fget, \fput)$, which in coend form is $\rep{[\id_S, \fget]}{\fput}$. Evaluating $\outside$ on this gives:
  \begin{align*}
    \outside(\rep{[\id_S, \fget]}{\fput}) = \fput [\id_S, \fget]
  \end{align*}
  so requiring $\outside(p) = \id_S$ is precisely the $\fget\fput$ law.

@@

  Additionally, the commutative diagram
  \[
    \begin{tikzcd}
      C \times A \ar[r, "i \times A"] \ar[d, "\fput (i \times A)", swap] & S \times A \ar[d, "\fput"] \\
      S \ar[r, equal] \ar[d, "{[j,\fget]}", swap] & S \ar[d, "{[\id_S, \fget]}"]  \\
      C \times A \ar[r, "i \times A", swap] & S \times A
    \end{tikzcd}
  \]
  is a witness that $\rep{\fput (i \times A)}{[j,\fget] } = \rep{\fput}{[\id_S, \fget] }$ as elements of $\Optic_\M(S, A)$.
\end{proof}

@@

Just as optics for $\times$ correspond to a pair of maps $\fget : S \to A$ and $\fput : S \times A \to S$, optics for $\sqcup$ correspond to pairs of maps $\freview : A \to S$ and $\fmatching : S \to S \sqcup A$. These names are taken from the Haskell \lenslib{} library.
\begin{align*}
  \Optic_\sqcup((S, S'), (A, A')) &= \int^{M \in \C} \C(S, M \sqcup A) \times \C(M \sqcup A', S') \\
                                  &\cong \int^{M \in \C} \C(S, M \sqcup A) \times \C(M, S') \times \C(A, S') && \text{(universal property of coproduct)} \\
                                  &\cong \C(S, S' \sqcup A) \times \C(A', S') && \text{(Yoneda reduction)}
\end{align*}
If we are given a prism $\rep{l}{r } : (S, S') \hto (A, A')$ then associated $\freview$ and $\fmatching$ morphisms are given by $\freview = r \inr$ and $\fmatching = (r\inl \sqcup A)l$

@@

\begin{proposition}
  An iso $\rep{l}{r}$ is lawful iff (as expected) $l$ and $r$ are mutual inverses.
\end{proposition}
\begin{proof}
  \[ \int^{M_1, M_2 \in \M} \C(S, M_1 A) \times \C(M_1 A, M_2 A) \times \C(M_2 A, S) \]
  specialises in this case to just
  \[ \C(S, A) \times \C(A, A) \times \C(A, S) \]
  The condition $\outside(\rep{l}{r }) = \id_S$ is the claim that $rl = \id_S$, and $\once(\rep{l}{r }) = (l, \id_A, r)$ is equal to $\twice(\rep{l}{r }) = (l, lr, r)$ iff $lr = \id_A$.
\end{proof}

@@

Suppose $\C$ is cartesian closed and let $\Strong_\C$ be the category of \emph{strong functors}\cite{StrongFunctors} on $\C$. Then, again, $\Strong_\C$ acts on $\C$ by evaluation. We leave it to the reader to verify there is a natural isomorphism \[\C(A, FB) \cong \Strong_\C(A \times \homC(B, -), F),\] which we can use to describe optics for this action as elements of $\C(\homC(A, A'), \homC(S,S'))$.
\end{remark}

@@

\subsection{Traversals}
Let $\C = \Set$. \todo{define applicative/traversable functors}

\begin{definition}
The category $\Traversal_\C$ of traversals in $\C$ is the category of lawful optics for the action of $\Traversable_\C$ on $\C$ given by evaluation. (Yes, the names $\Traversal$ vs\@. $\Traversable$ are confusing.)
\end{definition}

@@

\begin{proposition}[{\cite[Theorem 4.10, Proposition 5.4]{SecondOrderFunctionals}}]
Traversable structures on a functor $T : \Set \to \Set$ correspond to parameterised coalgebra structures
\begin{align*}
t_{A, B} : TA \to UR^*_{A, B}(T B)
\end{align*} 
where $UR^*_{X,Y}$ is the parameterised comonad
\begin{align*}
UR^*_{X, Y} Z = \Sigma_{n\in \bN} X^n \times \homC(Y^n,Z)
\end{align*}
Moreover, this correspondence forms an isomorphism of categories between $\Traversable$ and the Eilenberg-Moore category of coalgebras for $UR^*_{-, -}$, which we denote $\C_T$. \qed
\end{proposition}

@@

\begin{remark}
In~\cite[Section 2.3]{ProfunctorOptics}, it is claimed that a traversal $S \hto A$ exhibits an isomorphism \[S \cong \Sigma_n A^n \times \homC(A^n,S).\] This appears to be a misreading of~\cite[Proposition 5.4]{SecondOrderFunctionals}.
\end{remark}

@@

If we have a monoidal category $\M$ that acts on two different categories $\C_1$ and $\C_2$, we may well form the `optic' category:
\begin{align*}
  \Optic_\M((S, S'), (A, A')) := \int^{M \in \M} \C_1(S, MA) \times \C_2(MA', S')
\end{align*}
We lose, of course, the ability to talk about lawful optics. To encode \mintinline{haskell}{Getter}s or \mintinline{haskell}{Review}s, we then set $\C_2$ or $\C_1$ to be the codiscrete category $\C_\mathsf{codisc}$ with the same set of objects as $\C$, so that all homsets are the one element set. In this way we indeed get categories $\Getter$ and $\Review$ that are isomorphic to $\C$ and $\C^\op$. The functor $\C \to \C_\mathsf{codisc}$ induces functors $\Lens \to \Getter$ and $\Prism \to \Review$.

>>>>

\newcommand{\Act}{\mathbf{Act}}
\newcommand{\Optic}{\mathbf{Optic}}
\newcommand{\Lawful}{\mathbf{Lawful}}
%\newcommand{\SemiOptic}{\mathbf{SemiOptic}}
\newcommand{\Lens}{\mathbf{Lens}}
\newcommand{\Prism}{\mathbf{Prism}}
\newcommand{\Setter}{\mathbf{Setter}}
\newcommand{\Traversal}{\mathbf{Traversal}}
\newcommand{\Getter}{\mathbf{Getter}}
\newcommand{\Review}{\mathbf{Review}}
\newcommand{\Fold}{\mathbf{Fold}}

@@

\begin{lemma}[Ninja Yoneda Lemma/Yoneda Reduction]
For every functor $K : \C^\op \to \Set$ and $H : \C \to \Set$, we have the following natural isomorphisms:
\begin{align*}
K &\cong \int^C KC \times \C(-,C) &
K &\cong \int_C \Set(\C(C,-), KC) \\
H &\cong \int^C HC \times \C(C,-)  &
H &\cong \int_C \Set(\C(-,C), HC)
\end{align*}
where the components of the isomorphisms at $X$ are given by inclusion with the identity morphism $\C(X, X)$ on the left, and evaluation at the identity morphism on the right.
\qed
\end{lemma}

@@

Because this coend takes place in $\Set$, we can use Lemma \ref{lemma:calculate-coend} to describe $\Optic_\otimes((S, S'), (A, A'))$ explicitly. It is the set of pairs $(l, r)$, where $l : S \to M \otimes A$ and $r : M \otimes A' \to S'$, quotiented by the relation
\begin{align*}
  ((f \otimes A) l, r) \sim (l, r (f \otimes A'))
\end{align*}
for any $l : S \to M \otimes A$, $r : N \otimes A' \to S'$ and $f : M \to N$.

@@

  To check extranaturality of $\varepsilon$, suppose we have a dualisable optic $\iota(f, g) : (S, S') \hto (T, T')$, so $f : S \to T$ and $g : T' \to S'$. Luckily here all the switching in the definitions cancel out! Extranaturality is witnessed by the equality of the string diagrams:
  \begin{center}
    \input{diagrams/counit-extranatural-left.tikz}
    \qquad \raisebox{1.5cm}{$=$} \qquad
    \input{diagrams/counit-extranatural-right.tikz}
  \end{center}
  For symmetry:
  \begin{center}
    \input{diagrams/counit-symmetry-left.tikz}
    \qquad \raisebox{1.5cm}{$=$} \qquad
    \input{diagrams/counit-symmetry-right.tikz}
  \end{center}
  And for monoidality there is essentially nothing to do in the graphical calculus.
  \begin{center}
    \input{diagrams/counit-monoidal-left.tikz}
    \qquad \raisebox{2cm}{$=$} \qquad
    \input{diagrams/counit-monoidal-right.tikz}
  \end{center}

@@

  And it also preserves the counits:
  \begin{align*}
    F(\varepsilon_{(S, S')})
    &= F(c_{S \otimes S'}) \\
    &= F(\rep{\rho_S^{-1}}{\rho_{S'}}) \\
    &= \rep{\phi^{-1}_{S,I} (F \rho_S^{-1})}{(F \rho_{S'}) \phi_{S',I}} \\
    &= \rep{(FS \otimes \phi_I^{-1}) \phi^{-1}_{S,I} (F \rho_S^{-1})}{(F \rho_{S'}) \phi_{S',I} (FS \otimes \phi_I)} \\
    &= \rep{\rho_{FS}^{-1}}{\rho_{FS'}} \\
    &= \varepsilon_{(FS, FS')}
  \end{align*}

@@

\begin{definition}
Given two categories equipped with monoidal actions $(\M, \C)$ and $(\N, D)$, a \emph{morphism of actions} is a monoidal functor $F^\bullet : \M \to N$ and a functor $F : \C \to \D$ that commutes with the actions, in the sense that there exist natural isomorphisms
  \begin{align*}
  \phi &: F(I A) \to (F^\bullet I)(FA) \\
  \phi_{M,A} &: F(MA) \to (F^\bullet M)(F A)
  \end{align*}
satisfying conditions analogous to those for a monoidal functor.
\end{definition}

@@

  Suppose we have two lawful optics $\rep{l}{r} : R \hto S$ and $\rep{l'}{r'} : S \hto A$, where $l : R \to MS$ and $l' : S \to NA$. We must show that $\rep{(Ml')l}{r (Mr')}$ is also a lawful optic. The first equation is straightforward:
  \begin{align*}
    \outside(\rep{l; (Ml')}{(Mr') ; r})
    &= l ; Ml' ; Mr' ; r \\
    &= l ; M(l'r') ; r \\
    &= l ; M\id_{NA} ; r \\
    &= l ; r \\
    &= \id_S
  \end{align*}

  For the second, we must show
  \[ \repthree{ l;(Ml')}{(Mr'); r;l;(Ml')}{(Mr') ; r} = \repthree{l;(Ml')}{\id_{MNA}}{(Mr') ; r}. \]

@@

  So consider the generating relation $\repthree{ l;\phi_S}{c}{r} = \repthree{l}{\phi_S ; c}{r}$ in \[\int^{M_1, M_2 \in \M} \C(R, M_1 S) \times \C(M_1 S, M_2 S) \times \C(M_2 S, R).\]
  This implies that
  \begin{align*}
    \repthree{l;\phi_S;(Ml')}{(Mr'); c ;(Ml')}{(Mr') ; r}
    &= \repthree{l;(Ml');\phi_{NA}}{(Mr'); c ;(Ml')}{(Mr') ; r} && \text{(functoriality)} \\
    &= \repthree{l;(Ml')}{\phi_{NA};(Mr'); c ;(Ml')}{(Mr') ; r} && \text{(coend relation)} \\
    &= \repthree{l;(Ml')}{(Mr');\phi_{S}; c ;(Ml')}{(Mr') ; r} && \text{(functoriality)}
  \end{align*}
  And similarly for the other generating relation, $\repthree{l}{c;\phi_S}{r} = \repthree{l}{c}{\phi_S; r}$.

  Therefore, the chain of relations proving $\repthree{l}{r;l}{r} = \repthree{l}{\id_{MA}}{r}$ can be replicated to show
  \begin{align*}
    \repthree{l;(Ml')}{(Mr');r;l;(Ml')}{(Mr') ; r} &= \repthree{l;(Ml')}{(Mr');\id_{MA};(Ml')}{(Mr') ; r},
  \end{align*}
  which is equal to $\repthree{l;(Ml')}{M(r'l')}{(Mr') ; r}$.

  Similarly, a generating relation $\repthree{l';\psi_A}{c'}{r'} = \repthree{l'}{\psi_A ; c'}{r'}$ in \[\int^{N_1, N_2 \in \M} \C(S, N_1 A) \times \C(N_1 A, N_2 A) \times \C(M_2 A, S)\]
  implies that
  \begin{align*}
    \repthree{l;M(l';\psi_A)}{Mc'}{(Mr') ; r}
    &= \repthree{l;Ml';M\psi_A}{Mc'}{(Mr') ; r} \\
    &= \repthree{l;Ml'}{M\psi_A ; Mc'}{(Mr') ; r} \\
    &= \repthree{l;Ml'}{M(\psi_A;c')}{(Mr') ; r}
  \end{align*}
  So again we can replicate the chain of relations proving $\repthree{l'}{r';l'}{r'} = \repthree{l'}{\id_{MA}}{r'}$ to show that
  \[\repthree{l;(Ml')}{M(r'l')}{(Mr') ; r} = \repthree{l;(Ml')}{\id_{MNA}}{(Mr') ; r} \]

@@

\begin{proposition}
  Suppose $F : (\M, \C) \to (\N, \D)$ is a morphism of actions. Then $\Optic(F) : \Optic_\M \to \Optic_\N$ restricts to a functor $\Lawful_\M \to \Lawful_\N$.
\end{proposition}
\begin{proof}
  If $p = \rep{l}{r}$ is lawful, then verifying the first equation is easy:
  \begin{align*}
  \outside(\Optic(F)(\rep{l}{r}))
  &= \outside(\rep{(Fl);\phi^{-1}_{M,A}}{\phi_{M,A'};(Fr)}) \\
  &= (Fl);\phi^{-1}_{M,A};\phi_{M,A'};(Fr)\\
  &= (Fl);(Fr)\\
  &= \id_{FS}
  \end{align*}
  where $\phi_{M,A'} : (F^\bullet M)(FA) \to F(MA)$ is the structure map for $F$ commuting with the actions.

@@

\begin{proposition}
  Let $\rep{l}{r} : S \hto A$ be an optic. If $l;r = \id_S$ and $r;l = \phi A$ for some $\phi : M \to M$ in $\M$, then $\rep{l}{r}$ is lawful.
\end{proposition}
\begin{proof}
  The $\outside(\rep{l}{r}) = l;r = \id_S$ is exactly the first condition. And for the second, we verify:
  \begin{align*}
    \twice(\rep{l}{r})
    &= \repthree{l}{r;l}{r} \\
    &= \repthree{l}{\phi A}{r} && \text{($r;l = \phi A$)}\\
    &= \repthree{l (\phi A)}{\id_{MA}}{r} && \text{(coend relation)} \\
    &= \repthree{l;r;l}{\id_{MA}}{r} && \text{($r;l = \phi A$ again)}\\
    &= \repthree{l}{\id_{MA}}{r}  && \text{($l;r = \id_S$)}\\
    &= \once(\rep{l}{r})
  \end{align*}
\end{proof}

@@

Let $\inside : \Optic_\M(S, A) \to \int^{M \in \M} \C(M A, M A)$ be the map induced by $\inside(\rep{l}{r}) = \langle r ; l \rangle$. We might ask that instead of requiring $r;l = \phi A$ exactly, we have $\langle r ; l \rangle = \langle \phi A \rangle$ in $\int^{M \in \M} \C(M A, M A)$. In fact, this is equivalent:

\begin{proposition}\label{prop-onthenose}
  Suppose $p : S \hto A$ satisfies $\outside(p) = \id_S$ and $\inside(p) = \langle \phi A \rangle$. Then there exists a representative $\rep{l}{r}$ such that $l ; r = \psi A$ on the nose for some (possibly different) $\psi : M \to M$.
\end{proposition}
\begin{proof}
  The generating relation for $\int^{M \in \M} \C(M A, M A)$ is
  \[ \langle f \phi_A \rangle = \langle \phi_A f \rangle \]
  whenever $f : N A \to M A$ and $\phi : M \to N$. This relation $f \phi_A \rightsquigarrow \phi_A f$ is not likely to be symmetric or transitive in general.

@@

  We can now inductively apply the above argument to the shorter chain \[l'r' = u_1^2 \leftrightsquigarrow \dots \leftrightsquigarrow u_n^2 \leftrightsquigarrow (\psi A)^2 = \psi^2 A,\] obtained by squaring each morphism in the original chain, until we are left with a representative $\rep{l^*}{r^*}$ such that $r^*;l^* = \psi^N_A$, for some $N>0$. This pair $\rep{l^*}{r^*}$ is the required representative.
\end{proof}

@@

This last set is the set of concrete lenses $\conc_\times((S, S'), (A, A'))$. Explicitly this isomorphism states that, given an optic $\rep{l}{r} : (S, S') \hto (A, A')$, the corresponding element of $\conc_\times((S, S'), (A, A'))$ is the pair $(\fget, \fput)$, where $\fget = \pi_2 l$ and $\fput = r (\pi_1 l \times A)$. In the other direction, given $(\fget, \fput)$, the corresponding optic has representative $\rep{[\id_S, \fget]}{\fput}$.

@@

\begin{proposition}\label{prop-OpticImpliesLensLaws}
  A concrete lens given by $\fget$ and $\fput$ is lawful (in our sense) if and it obeys the three lens laws.
\end{proposition}
\begin{proof}
  We begin by giving
  \[ \int^{M_1, M_2 \in \C} \C(S, M_1 \times A) \times \C(M_1 \times A, M_2 \times A) \times \C(M_2 \times A, S)\]
  the same treatment as $\Optic_\times(S, A)$. By the same sorts of arguments as above \todo{is this worth writing out?}, this is isomorphic to
  \[\conctwice_\times(S, A) := \C(S, A) \times \C(S \times A, A) \times \C(S \times A \times A, S)\]
  with the isomorphism given by
  \begin{align*}
    \Phi(\repthree{l}{c}{r}) = (&\pi_2 l, \\
                                   &\pi_2 c (\pi_1 l \times A), \\
                                   &r(\pi_1 c (\pi_1 l \times A) \times A))
  \end{align*}
  Now suppose we are given a lens $(\fget, \fput)$, which in coend form is $\rep{[\id_S, \fget]}{\fput}$. Evaluating $\outside$ on this gives:
  \begin{align*}
    \outside(\rep{[\id_S, \fget]}{\fput}) = \fput [\id_S, \fget]
  \end{align*}
  so requiring $\outside(p) = \id_S$ is precisely the $\fget\fput$ law.

@@

  Additionally, the commutative diagram
  \[
    \begin{tikzcd}
      C \times A \ar[r, "i \times A"] \ar[d, "\fput (i \times A)", swap] & S \times A \ar[d, "\fput"] \\
      S \ar[r, equal] \ar[d, "{[j,\fget]}", swap] & S \ar[d, "{[\id_S, \fget]}"]  \\
      C \times A \ar[r, "i \times A", swap] & S \times A
    \end{tikzcd}
  \]
  is a witness that $\rep{\fput (i \times A)}{[j,\fget]} = \rep{\fput}{[\id_S, \fget]}$ as elements of $\Optic_\M(S, A)$.
\end{proof}

@@

Just as optics for $\times$ correspond to a pair of maps $\fget : S \to A$ and $\fput : S \times A \to S$, optics for $\sqcup$ correspond to pairs of maps $\freview : A \to S$ and $\fmatching : S \to S \sqcup A$. These names are taken from the Haskell \lenslib{} library.
\begin{align*}
  \Optic_\sqcup((S, S'), (A, A')) &= \int^{M \in \C} \C(S, M \sqcup A) \times \C(M \sqcup A', S') \\
                                  &\cong \int^{M \in \C} \C(S, M \sqcup A) \times \C(M, S') \times \C(A, S') && \text{(universal property of coproduct)} \\
                                  &\cong \C(S, S' \sqcup A) \times \C(A', S') && \text{(Yoneda reduction)}
\end{align*}
If we are given a prism $\rep{l}{r} : (S, S') \hto (A, A')$ then associated $\freview$ and $\fmatching$ morphisms are given by $\freview = r \inr$ and $\fmatching = (r\inl \sqcup A)l$

@@

\begin{proposition}
  An iso $\rep{l}{r}$ is lawful iff (as expected) $l$ and $r$ are mutual inverses.
\end{proposition}
\begin{proof}
  \[ \int^{M_1, M_2 \in \M} \C(S, M_1 A) \times \C(M_1 A, M_2 A) \times \C(M_2 A, S) \]
  specialises in this case to just
  \[ \C(S, A) \times \C(A, A) \times \C(A, S) \]
  The condition $\outside(\rep{l}{r}) = \id_S$ is the claim that $rl = \id_S$, and $\once(\rep{l}{r}) = (l, \id_A, r)$ is equal to $\twice(\rep{l}{r}) = (l, lr, r)$ iff $lr = \id_A$.
\end{proof}

@@

Suppose $\C$ is cartesian closed and let $\Strong_\C$ be the category of \emph{strong functors}\cite{StrongFunctors} on $\C$. Then, again, $\Strong_\C$ acts on $\C$ by evaluation. We leave it to the reader to verify there is a natural isomorphism \[\C(S, FA) \cong \Strong_\C(\homC(A, -) \times S, F),\] which we can use to describe optics for this action as elements of $\C(\homC(A, A'), \homC(S,S'))$.
\end{remark}

@@

\subsection{Traversals}
In this section we work in the case $\C = \Set$. \todo{define applicative/traversable functors}

\begin{definition}
The category $\Traversal$ of traversals is the category of lawful optics for the action of $\Traversable$ on $\Set$ given by evaluation. (Yes, the names $\Traversal$ vs\@. $\Traversable$ are confusing.)
\end{definition}

@@

\begin{proposition}[{\cite[Theorem 4.10, Proposition 5.4]{SecondOrderFunctionals}}]
Traversable structures on a functor $T : \Set \to \Set$ correspond to parameterised coalgebra structures
\begin{align*}
t_{A, B} : TA \to UR^*_{A, B}(T B)
\end{align*}
where $UR^*_{X,Y}$ is the parameterised comonad
\begin{align*}
UR^*_{X, Y} Z = \Sigma_{n\in \bN} X^n \times \Set(Y^n,Z)
\end{align*}
Moreover, this correspondence forms an isomorphism of categories between $\Traversable$ and the Eilenberg-Moore category of coalgebras for $UR^*_{-, -}$, which we denote $\E$. \qed
\end{proposition}

@@

\begin{remark}
In~\cite[Section 2.3]{ProfunctorOptics}, it is claimed that a traversal $S \hto A$ exhibits an isomorphism \[S \cong \Sigma_n A^n \times \Set(A^n,S).\] This appears to be a misreading of~\cite[Proposition 5.4]{SecondOrderFunctionals}.
\end{remark}

@@

If we have a monoidal category $\M$ that acts on two different categories $\C_1$ and $\C_2$, we may form the `optic' category:
\begin{align*}
  \Optic_\M((S, S'), (A, A')) := \int^{M \in \M} \C_1(S, MA) \times \C_2(MA', S')
\end{align*}
We lose, of course, the ability to talk about lawful optics. To encode \mintinline{haskell}{Getter}s or \mintinline{haskell}{Review}s, we then set $\C_2$ or $\C_1$ to be the codiscrete category $\C_\mathsf{codisc}$ with the same set of objects as $\C$, so that all homsets are the one element set. In this way we indeed get a category $\Getter$ that is isomorphic to $\C$.
\begin{align*}
\int^{M \in \M} \C(S, M \times A) \times \C_\mathsf{codisc}(M \times A', S')
&\cong \int^{M \in \M} \C(S, M \times A) \\
&\cong \C(S, 1 \times A) \\
&\cong \C(S, A)
\end{align*}
This works for $\Review$ in the same way, putting $\C_\mathsf{codisc}$ on the left. The functor $\C \to \C_\mathsf{codisc}$ induces functors $\Lens \to \Getter$ and $\Prism \to \Review$.
